---
title: "LEMA Irrigation Trends"
author: "Jill Deines"
date: "September 25, 2017"
output: 
  html_document:
    toc: yes
---

Goal: Investigate trends in irrigated area over time in the Sheridan 6 LEMA compared to surrounding areas

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path='figure/00.1_irrTrends_aim/')
```

**R Packages Needed**

```{r packages, warning=FALSE, message=FALSE}
library(tidyverse) # for ggplot2, tidyr, dplyr
```

## Irrigation Over Time

### load RS estimates of irrigated area
Counts of each class type were tallied for several administrative boundaries in GEE

```{r loadGee}
cellArea <- 30*30 # m^2

regionsDir <- 'C:/Users/deinesji/Google Drive/GEE_AIM-RRB/GEE_LEMA/GEE_LEMA_tableExports'

# regional areas -----------------------------------------------
# get filenames
files <- list.files(regionsDir, pattern="*_AIMRRB_irrigatedArea_majorBoundaries.csv")

# read in to a single dataframe
loadStats <- function(fileList){
  datOut = as.data.frame(do.call(rbind, lapply(fileList, function(x) {
              csv1 <- read.csv(paste0(regionsDir,'/',x))
              csv1$year <- as.numeric(substr(x, start=1,stop=4))
              return(csv1)
           })))  
  # convert cell count to area
  datOut$irrigated_m2 <- datOut$X1 * cellArea
  datOut$irrigated_km2 <- datOut$irrigated_m2 / 1000000
  return(datOut)
}

all1 = loadStats(files)

# reformat data
all <- all1[,c('masterid','year','irrigated_m2','irrigated_km2','pr_ann')]
names(all)[3:4] <- c('area_m2', 'area_km2')
```

### side note: pull out climate data
get a file of climate and area to export

```{r getclimate}
regionsWanted <- c('sheridan6','KS_GMD4','RRCA_mod_KS_main')
climate <- all1[all1$masterid %in% regionsWanted,
                c('masterid','year','irrigated_km2','pr_ann','pr_early',
                   'pr_grow','pdsi_ann','pdsi_grow')]
# export
outdir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/LEMAs/data/tabular'
write.csv(climate, paste0(outdir,'/irrArea_climate_from_AIMRRB.csv'),
          row.names=FALSE)
```


### Plot Preprocessing

Extract regions wanted and calculate cumulative added and lost irrigation. Set some overly? complicated plot parameters

```{r LEMAtrendsByLine}
# extract regions to plot
regionsWanted <- c('sheridan6','KS_GMD4','RRCA_mod_KS_main')
regions0 <- all[all$masterid %in% regionsWanted,]

# nice names
key <- data.frame(masterid = c('sheridan6','KS_GMD4','RRCA_mod_KS_main'),
                  region = c('Sheridan 6', 'GMD4', 'KS RRCA'))
regions <- merge(regions0, key)

# plots
ggplot(regions, aes(x = year, y = area_km2, color=region)) +
  geom_line() +
  ylab('Irrigated Area (km^2)') + xlab('') +
  theme_bw() + theme(legend.position = 'top',
                     legend.title = element_blank())
```

```{r LemaTrendsFacet, fig.width=4.5, fig.height = 4.5}
ggplot(regions, aes(x = year, y = area_km2)) +
  geom_line() +
  facet_wrap(~region, scales = "free_y", nrow = 3) +
  ylab('Irrigated Area (km^2)') + xlab('') +
  theme_bw() + theme(legend.position = 'top',
                     legend.title = element_blank())

ggplot(regions, aes(x = year, y = area_km2)) +
  geom_line() + geom_smooth() +
  facet_wrap(~region, scales = "free_y", nrow = 3) +
  ylab('Irrigated Area (km^2)') + xlab('') +
  theme_bw() + theme(legend.position = 'top',
                     legend.title = element_blank())

ggplot(regions, aes(x = year, y = area_km2)) +
  geom_line() + stat_smooth(method = 'lm') +
  facet_wrap(~region, scales = "free_y", nrow = 3) +
  ylab('Irrigated Area (km^2)') + xlab('') +
  theme_bw() + theme(legend.position = 'top',
                     legend.title = element_blank())
```


