---
title: "WIMAS Exploratory Analyses"
author: "Jill Deines"
date: "October 12, 2017"
output: 
  html_document:
    toc: yes
---

Goal: Exploratory analyses of the well data as they pertain to LEMA

Secondary goal: Get familiar with the `sf` package, a new vector GIS package

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path='figure/01.05_WIMASexploratory/',
                      cache = TRUE)
```

**R Packages Needed**

```{r packages, warning=FALSE, message=FALSE}
library(sf)
library(tidyverse)
library(xtable)
```

## Load data files
Includes formatted data from 01.05 and boundary shapefiles of interest

```{r loadData}
# directories
gisDir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/LEMAs/data/GIS/boundaries'
wellDir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/LEMAs/data/wellData/WIMAS_AnthonyCleaned_JillFormatted'

# well data
wimas <- read.csv(paste0(wellDir, '/WIMAS_1990-2016_gwPts_ksHpa.csv'))

# boundaries of interest
sheridan <- st_read(paste0(gisDir,'/Sheridan6_trace.shp'), quiet=TRUE)
sheridan$Id <- 'Sheridan'

gmds.ll <- st_read(paste0(gisDir, '/KS - GROUNDWATER_DISTRICTS.shp'),quiet = TRUE)
KSrrca <- st_read(paste0(gisDir,'/RRCA-KS-upperOnly_aea.shp'),quiet = TRUE)
```

Normalize projections: let's work in EPSG:5070 to match my GEE work.

Assuming well points are WGS84. That should be checked.

```{r reproject}
sheridan <- st_transform(sheridan, 5070)
gmds <- st_transform(gmds.ll, 5070)
KSrrca <- st_transform(KSrrca, 5070)

# visualize
plot(gmds[1], main='Some Regions', graticule = st_crs(gmds.ll), axes=TRUE)
plot(KSrrca, add=TRUE)
plot(sheridan, add=TRUE)

# spatialize well data
wimasLL <- st_as_sf(x = wimas,
                      coords = c('LONGITUDE','LATITUDE'),
                      crs = "+proj=longlat +datum=WGS84")
wimas.aea <- st_transform(wimasLL, 5070)

# subset 1 year of points for reasonable plotting
wimasOne <- wimas.aea[wimas.aea$year == 2000,]

# plot points
plot(gmds[1], main='WIMAS points', graticule = st_crs(gmds.ll), axes=TRUE)
plot(wimasOne[1], add=TRUE)
```

## Regional Trends

### Sheridan 6

* How many wells are there?
* Can you see the 20% reduction in pumping from the LEMA?

Note: for the moment I'm assume volume is in acre ft.

```{r sheridan1}
# extract points in sheridan
wellsLema <- st_join(wimas.aea, sheridan) %>% filter(Id == 'Sheridan')

# how many?
length(unique(wellsLema$PDIV_ID))

# plot wells
plot(sheridan, main = '222 Wells in Sheridan 6')
plot(wellsLema[wellsLema$year ==2000,1], add=TRUE, pch = 19)

lemakey <- data.frame(year = 1990:2016,
                      LEMA = c(rep('preLema',23), rep('lema',4)))

wellsLema2 <- merge(wellsLema, lemakey)

prepost <- wellsLema2 %>%
  group_by(LEMA) %>%
  summarize(meanVol = mean(volume, na.rm=T), 
            meanAcres = mean(acres, na.rm=T),
            meanDepth = mean(depth, na.rm=T))

longg <- gather(prepost, key = dataset, value = value, meanVol:meanAcres)

ggplot(longg, aes(x=LEMA, y = value)) + 
  geom_bar(stat='identity') + facet_wrap(~dataset)
```

A more nuanced approach: build a linear model to predict LEMA years, and then compare. The purpose of this is to account for precipitation and general trends in decreasing irrigation depths.

Check out trends

```{r sheridanTimeSeries, fig.width=4.5}
# plot all wells over time
wellslong <- gather_(wellsLema2, key = 'variable', value = 'value', 
                     c('acres','volume', 'depth'))

# see acres and volume over time, all wells
ggplot(wellslong, aes(x=year, y = value)) +
  geom_line(aes(group = PDIV_ID), lwd=.5) +
  facet_wrap(~variable, scales = 'free_y', nrow = 3) + 
  theme_bw() + ggtitle('Sheridan Area, Depth, & Volume over time')

# see acres and volume over time, all wells - boxplot!
ggplot(wellslong, aes(x=year, y = value)) +
  geom_boxplot(aes(group=year)) +
  facet_wrap(~variable, scales = 'free_y', nrow = 3) + 
  theme_bw() + ggtitle('Sheridan Area, Depth, & Volume over time')

# plot volume vs acres
ggplot(wellsLema2, aes(x=volume, y=acres, color=year)) +
  geom_point() + ggtitle('Sheridan volume vs area')

# depth over time
ggplot(wellsLema2, aes(x=year, y=depth)) +
  geom_boxplot(aes(group=year)) + ylab('depth (ft)')+
  ggtitle('Sheridan depth over time')


#  total over time
summed <- wellsLema2 %>%
  group_by(year) %>%
  summarize(totalVol = sum(volume, na.rm=T), totalAcres = sum(acres, na.rm=T)) %>%
  mutate(meandepth = totalVol/totalAcres)

ggplot(summed, aes(x=totalVol, y=totalAcres, color=year)) +
  geom_point() + ggtitle('Sheridan: Total volume vs total area')

sumLong <- gather(summed, key = variable, value =value, totalVol:meandepth)

ggplot(sumLong, aes(x=year, y = value)) +
  geom_line() +
  facet_wrap(~variable, scales = 'free_y', nrow = 3) + 
  theme_bw() + ggtitle('Sheridan Total Volume & Area over time')
```

Many PDIV_ID's seem to go to zero on occasion - how will that affect a prediction?

Maybe try mean depth ~ precip + year

```{r depthModel}

```

