---
title: "Crop Type Breakdowns"
author: "Jill Deines"
date: "November 6, 2017"
output: 
  html_document:
    toc: yes
---

Goal: Summarize dominant crops in the LEMA project study area (greater GMD4 region)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path='figure/01.50_CDL_summary/',
                      cache = TRUE)
```

**R Packages Needed**

```{r packages, warning=FALSE, message=FALSE}
library(tidyverse)
library(stringr)
```

### user variables

```{r uservars}
# cdl summaries by counties, sheridan
rawDir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/LEMAs/data/tabular/cdl_tables_lema'
fileNameSuffix <- '_CDL_GMD4Plus_Sheridan_area_m2.csv'

# cdlkey
cdlkeyfile <-'C:/Users/deinesji/Dropbox/1PhdJill/hpa/LEMAs/data/tabular/cdl_tables_lema/CDL_key_2014.csv'

# output data directory
outDir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/LEMAs/data/tabular'
outFnameCounty <- 'CDL_byCounty_GMD4pClip_2006-2016.csv'
outFnameSheridan <- 'CDL_SheridanTrace_2006-2016.csv'
```

# Load Data and Reformat to Tidy Dataframe
CDL summaries (area of each class by polygons within the GMD4 Plus boundary - counties and Sheridan 6) were exported from Google Earth Engine using script
in users/jdeines/default/lema

Note: 2015 and 2015 have extra cdl class columns: 251,252,253. In all national metadata files for 2008 - 2016, these values are all NA. So, dropping columns. Most come with a 0 column which is nonsense, removing so all match. 2013 has a class 62 that is NA for all others, for 'pasture/grass', but no classified cells for this class, so removing that as well.

```{r loadCDL}
# get filenames
files <- list.files(rawDir, pattern=paste0('*', fileNameSuffix))

# read in to a single dataframe, removing junk columns and first dummy row
dataAll <- as.data.frame(do.call(rbind, lapply(files, function(x) {
  csv1 <- read.csv(paste0(rawDir,'/',x))
  csv1 <- csv1[, -which(names(csv1)%in% c('system.index','.geo','X0','X62',
                                          'X251','X252','X253'))]
  csv1 <- csv1[-1,]    # remove dummy row
})))  

# add a crop type column using cdl key
cdlkey <- read_csv(cdlkeyfile) %>% 
  select(VALUE, CLASS_NAME) %>%
  rename(cdlCode = VALUE, cdlClassName = CLASS_NAME)

# convert to long data format, remove NA's, reformat cdlCodes, join with class name
dataLong <- gather(dataAll, key = cdlCode, value = area_m2, X1:X92) %>%
  filter(!is.na(area_m2)) %>%
  mutate(cdlCode = as.numeric(substr(cdlCode, start = 2, 
                                      stop = nchar(cdlCode))),
         area_km2 = area_m2 * 1e-6) %>%
  left_join(cdlkey)
```

## Split Dataset
To ensure no area is double counted, I split out the Sheridan 6 polygon from the county numbers. The `masterid` column is split into relevant variables

I'm pretty sure there are sometimes 2 entries in a year/crop class for Sheridan because it's a multipolygon. I condense these.

```{r splitData}
# Get the Sheridan dataset
sheridan <- dataLong %>% 
  filter(masterid == 'SheridanTrace') %>%
  group_by(Year, cdlCode, cdlClassName) %>%   # reduce split poly to 1 yearly num
  summarise(area_km2 = sum(area_km2)) %>%
  arrange(Year, cdlCode) 

write.csv(sheridan, row.names = FALSE, file = paste0(outDir,'/',outFnameSheridan))

# And the county dataset
gmd4.county <- dataLong %>%
  filter(masterid != 'SheridanTrace') %>%
  mutate(fips5 = str_extract(masterid,'[:digit:]+')) %>%
  arrange(Year, fips5)

write.csv(gmd4.county, row.names = FALSE, file = paste0(outDir,'/',outFnameCounty))
```

# Sheridan Stats
What are the CDL classes that cover 95% of the area?

```{r sheridan, fig.width = 5, fig.height = 3.5}
# crops ever planted
unique(sheridan$cdlClassName)

# mean acres by crop
cropMeans <- sheridan %>%
  group_by(cdlClassName) %>%
  summarise(meanArea_km2 = mean(area_km2)) %>%
  arrange(-meanArea_km2) %>%
  mutate(totalArea = sum(meanArea_km2),
         cumArea = cumsum(meanArea_km2),
         proportion = cumArea/totalArea)

# check it
head(cropMeans)

# keep top classes only
majorCrops <- cropMeans %>%
  filter(proportion < 0.96)

shortnames <- data.frame(cdlClassName = majorCrops$cdlClassName,
                         Commodity = c('Corn','Grass','W. Wheat','Fallow',
                                       'Sorghum','Soybeans'))
majorCrops <- majorCrops %>% 
  left_join(shortnames) %>%
  left_join(cdlkey)

# plot
position <- majorCrops$Commodity
ggplot(majorCrops, 
       aes(x=Commodity, y=meanArea_km2)) +
  geom_bar(stat='identity') + 
  ylab(expression(paste('Area (',km^2,')',sep=''))) + 
  scale_x_discrete(limits = position) +
  ggtitle('Sheridan 6, 2006-2016 mean') + 
  theme_bw() + theme(legend.title=element_blank(),
                     axis.text=element_text(size=10),
                     legend.text=element_text(size=10),
                     axis.title=element_text(size=11),
                     panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank())

majorCrops
```

# gmd4 stats
again, what covers 95% of the area?

NOTE: This is for the buffered study region, not strictly GMD4. I'll have to re-run the GEE script to get GMD4.

```{r gmd4, fig.width = 5, fig.height=3.5}
# crops ever planted
unique(gmd4.county$cdlClassName)

# mean acres by crop for the full gmd4 (ignore county bounds)
gmd4Means <- gmd4.county %>%
  group_by(Year, cdlClassName) %>%
  summarise(area_km2 = sum(area_km2)) %>%  # get total across all counties by year
  group_by(cdlClassName) %>%
  summarise(meanArea_km2 = mean(area_km2)) %>%
  arrange(-meanArea_km2) %>%
  mutate(totalArea = sum(meanArea_km2),
         cumArea = cumsum(meanArea_km2),
         proportion = cumArea/totalArea)

# check it
head(gmd4Means)

# keep top classes only
majorCrops2 <- gmd4Means %>%
  filter(proportion < 0.96)

shortnames2 <- data.frame(cdlClassName = majorCrops2$cdlClassName,
                         Commodity = c('Grass','W. Wheat','Corn','Fallow',
                                       'Sorghum'))
majorCrops2 <- majorCrops2 %>% 
  left_join(shortnames2) %>%
  left_join(cdlkey)

# plot
position <- majorCrops2$Commodity
ggplot(majorCrops2, 
       aes(x=Commodity, y=meanArea_km2)) +
  geom_bar(stat='identity') + 
  ylab(expression(paste('Area (',km^2,')',sep=''))) + 
  scale_x_discrete(limits = position) +
  ggtitle('GMD4 + buffer, 2006-2016 mean') + 
  theme_bw() + theme(legend.title=element_blank(),
                     axis.text=element_text(size=10),
                     legend.text=element_text(size=10),
                     axis.title=element_text(size=11),
                     panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank())
majorCrops2
```

