---
title: "Salus Weather Generator - GRIDMET"
author: "Jill Deines"
date: "December 7, 2017"
output: 
  html_document:
    toc: yes
---

Goal: Create SALUS weather input (.wdb.xml) using GRIDMET data obtained from Google Earth Engine, since NLDAS was wonky.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path='figure/02.06_SALUS_weather_gridmet/',
                      cache = TRUE)
```

**R Packages Needed**

```{r packages, warning=FALSE, message=FALSE}
# to get nldas polygons and centroids for region of interest
library(tidyverse)
```

# Input Data from GEE

Note: Gridmet data accessed with Google Earth Engine script users/jdeines/default/LEMA/02.02_Get_ClimateData run on 12/7/17

GRIDMET data was sampled to polygons of the NLDAS grid that include centroid coordinates and NLDAS cell elevation, created in 2.05_SALUS_weather.Rmd. Here's that code for posterity

```{js eval=FALSE}
// get gridmet for salus
// december 7 2017

// goal: sample gridmet at the nldas grid resolution for teh salus input file. variables 
// retrieved include year, doy, srad, tmax, tmin

// boundary extent
var polygrid = ee.FeatureCollection('users/jdeines/vector/RRB/NLDAS_clipped_gmd4p');
print(polygrid);

// set years
var startYear = 2000;
var endYear = 2017;

// collection info
var climateName = 'IDAHO_EPSCOR/GRIDMET'
var variables = ['pr','tmmn','tmmx','srad']

var tempAdjust = -273.15;
var sradConversion = 0.0864;

// export info
var exportFolder = 'lema_gridmet'
var filename = 'Gridmet_gmd4p_nldasPolys'


// end user vars --------------------------------------------------------

var years = []
for (var i = startYear; i <= endYear; i++) {
    years.push(i)
}
var startDate = startYear + '-01-01';
var endDate = endYear + '-12-31'

// get mean annual precip by year 
//var gridSum = years.map(function(year){
var year = 1999;  
  var startDate = year + '-01-01';
  var endDate = year + '-12-31'

  var gridmet = ee.ImageCollection(climateName)
      .select(variables)
      .filterDate(startDate, endDate);
      
  var gridmetAdjusted = gridmet.map(function(image){
    var precip = image.select(['pr']);
    var temps = image.select(['tmmn','tmmx']).add(tempAdjust);
    var srad = image.select(['srad']).multiply(sradConversion);
    return ee.Image.cat(precip, temps,srad).copyProperties(image)    
  });
  print(gridmetAdjusted)
      
  // get mean by polygon
  var triplets = gridmetAdjusted.map(function(image) {
  return image.reduceRegions({
    collection: polygrid,
    reducer: ee.Reducer.mean(),
    scale: 2000
  }).filter(ee.Filter.neq('mean', null))
    .map(function(f) {
      return f.set('day', image.get('system:index'));
  });
}).flatten();

print(triplets.limit(1));

// drop .geo contents
var tripletsOut = triplets.select(['.*'], null, false);


Export.table.toDrive({
 collection: tripletsOut,
  description: filename,
  folder: exportFolder,
 fileFormat: filename
});
```

# User Variables 

```{r userVars}
# gridmet folder
gridDir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/LEMAs/data/salus/weather/gridmet_Jill'
gridDataName <- 'Gridmet_gmd4p_nldasPolys.csv'

# NLDAS date bounds
startDate <- '2000-01-01'
endDate <- '2017-10-31'

# output subfolder
outXmlDir <- 'gridmet_nldas_grid_20000101_20171201'
```

**END USER dEFINED VARIABLES**

## Functions
Functions to write out the wdb.xml top and bottom matter.

```{r funs}
# outFile <- paste0(gridDir, '/', outXmlDir, '/test.wdb.xml')
# stationID <- '1100'
# lat <- 45.3
# lon <- 113.4
# elev <- 938.12

write_wdb_top <- function(outFile,stationID, lat, lon, elev){
  cat(paste0('<WDB>\n  <Stations StationID="', stationID, '" Station_Name="', stationID, 
             '" Place_Name="" Lat="', lat, '" Long="', lon, '" Elev="', elev,
             '" Tav="" Amp="" Tmht="2" Wmht="">\n',
             '    <Storm_Intensity>\n','    </Storm_Intensity>\n',
             '    <Hourly_Rainfall>\n','    </Hourly_Rainfall>\n',
             '    <Weather Columns="Year,DOY,SRAD,Tmax,Tmin,Rain,DewP,Wind,PAR">\n'),
      file=outFile)
}

#write_wdb_top(outFile, stationID, lat, lon, elev)

write_wdb_bottom <- function(outFile){
  cat(paste0('    </Weather>\n  </Stations>\n  <Version_Control>\n',
             '    <Version Number="1">\n',
             '      <ReleaseDate>11/26/1999 12:00:00 AM</ReleaseDate>\n',
             '      <Notes>Initial release, beta version.</Notes>\n',
             '    </Version>\n    <Version Number="1.1">\n',
             '      <ReleaseDate>11/27/1999 12:00:00 AM</ReleaseDate>\n',
             '      <Notes>testing...</Notes>\n    </Version>\n',
             '    <Version Number="X_1.1">\n      <ReleaseDate>03/20/2014</ReleaseDate>\n',
             '      <Notes>Automatically converted from Access DB by',
             ' salus_access2xmlprogram.</Notes>\n',
             '    </Version>\n  </Version_Control>\n</WDB>'),
      file=outFile, append=TRUE)
}

#write_wdb_bottom(outFile)
```

## Load data
And do some cleaning

* precip can't be negative (very very small negatives from encodings, i think i read on EE forum)

```{r loaddata}
# load data and keep key columns needed in wdb.xml
gridData <- read_csv(paste0(gridDir, '/', gridDataName)) %>%
  dplyr::select(c(gridCll, Lat, Lon, meanElv, day, srad, tmmx, tmmn, pr))

# clean precip data - no negatives
sum(gridData$pr < 0)
head(gridData[gridData$pr < 0,'pr'])
gridData[gridData$pr < 0,'pr'] <- 0
sum(gridData$pr < 0)

# check for other missing data
sum(is.na(gridData$tmmn))
sum(is.na(gridData$tmmx))
sum(is.na(gridData$srad))
sum(is.na(gridData$pr))

# round data to 2 decimals and make year and doy columns
gridData2 <- gridData %>%
  mutate(srad = round(srad, 2),
         tmmx = round(tmmx, 2),
         tmmn = round(tmmn, 2),
         pr = round(pr, 2),
         year = as.numeric(substr(day, start =1, stop = 4))) %>%
  group_by(gridCll, year) %>%
  arrange(gridCll,day) %>%
  mutate(DOY = 1:n()) 
```

## Write Wdb's!
A separate wdb.xml for each grid cell

```{r write_em, eval=FALSE}
# split   and write out
junk <- by(gridData2, gridData2$gridCll, function(x){
  filename <- paste0(gridDir, '/', outXmlDir, '/', x$gridCll[1], '.wdb.xml')
  # write top matter
  write_wdb_top(filename, x$gridCll[1], x$Lat[1], x$Lon[1], x$meanElv[1])
  # write data table
  outdf <- x[,c('year','DOY','srad','tmmx','tmmn','pr')]
  outdf$dewp <- NA
  outdf$wind <- NA
  outdf$par <- NA
  write.table(outdf, filename, append=TRUE, na = '', 
              sep = ',', row.names=FALSE, col.names=FALSE)
  # write the bottom matter
  write_wdb_bottom(filename)
})
```



