---
title: "SALUS Conceptual Model"
author: "Jill Deines"
date: "October 31, 2017"
output: 
  html_document:
    toc: yes
---

Goal: Look at input datasets and gauge number of SALUS models needed

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path='figure/01.05_WIMASexploratory/',
                      cache = TRUE)
```

**R Packages Needed**

```{r packages, warning=FALSE, message=FALSE}
library(tidyverse)
library(rgdal)
library(raster)
library(ncdf4)
library(sf)
```

# LEMA boundaries
Load boundaries of interest

```{r loadBounds}
gisDir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/LEMAs/data/GIS/boundaries'

lema <- readOGR(gisDir, 'Sheridan6_trace', verbose = FALSE)
gmd4.ll <- readOGR(paste0(gisDir,'/KS_GMD4.kml'), verbose=FALSE)
counties <- readOGR(gisDir, 'Counties_RRCAextent_Clip', verbose = FALSE)

# reproject
wgs84proj <- "+proj=longlat +datum=WGS84"
lema.ll <- spTransform(lema, CRS(wgs84proj))
```


# Weather Grid Options

* NLDAS: 1/8 degree (...12.5km?)
* GRIDMET: ~ 4km (1/24 degree)

NLDAS template copied from script in RRB_Rcode project. GRIDMET information acquired from https://climate.northwestknowledge.net/METDATA/. I downloaded the elevation dataset provided on the same 4 km grid in order to extract the spatial grid.

### extract gridmet grid

```{r getGridmetGrid, eval=FALSE}
# load elevation grid
gridmetDir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/LEMAs/data/weather/gridmet'
elev <- raster(paste0(gridmetDir,'/metdata_elevationdata.nc'))

# check out parameters
elev
```

## Compare Grids

### specify grid templates with cell number for future ID

```{r makeGrids}
# create a NLDAS raster grid template (full US) using grid specs
adjust <- 0.0625 # half of the 1/8 degree grid
nldasTemplate <- raster(nrow = 224, ncol = 464, crs = "+proj=longlat +datum=WGS84",
                     extent(-124.9375 - adjust,     #ymin
                            -67.0625 + adjust,      #ymax
                            25.0625 - adjust,       #xmin
                            52.9375 + adjust))      #xmax
# assign grid ID's
nldasTemplate[] <- 1:ncell(nldasTemplate)

# create a Gridmet grid template
metTemplate <- raster(nrow = 585, ncol = 1386, crs = "+proj=longlat +datum=WGS84",
                     extent(-124.793,     #ymin
                            -67.0439,      #ymax
                            25.04224,       #xmin
                            49.41686))      #xmax
# assign grid ID's
metTemplate[] <- 1:ncell(metTemplate)
```

### Clip grids to greater GMD4 area

```{r clipToGMD}
# crop to gmd
metCrop <- crop(metTemplate, gmd4.ll)
nldasCrop <- crop(nldasTemplate, gmd4.ll)

# polygonize
nldas.poly <- rasterToPolygons(nldasCrop)
met.poly <- rasterToPolygons(metCrop)

# view
par(mar = c(0,0,0,0))
plot(met.poly, lwd = .5)
plot(nldas.poly, border = 'blue', add = TRUE, lwd = 2)
plot(gmd4.ll, border = 'magenta', lwd = 2, add=TRUE)
plot(lema.ll, border = 'red', lwd = 2, add=TRUE)
```

# Check out county lines

```{r countyPlot}
par(mar = c(0,0,0,0))
plot(gmd4.ll, col = 'gray90', border = 'gray70')
plot(counties, add = TRUE)
plot(lema.ll, add = TRUE, col = 'indianred')
```

Holy moly, that's a small region! I'm thinking about doing those three counties fully within GMD4 in order to be able to validate yield estimates.

What are those counties?

```{r pickcounties}
# add a buffer to GMD4
gmdbuff <- st_as_sf(gmd4.ll) %>% 
  st_transform(5070) %>%
  st_buffer(5000) %>%
  st_transform(4326)

plot(gmdbuff[1], main = 'GMD4 5 km buffer')
plot(gmd4.ll[1], add=TRUE, col = 'gray80')

# intersect counties and GMD4
gmdCounties <- st_as_sf(counties) %>% 
  st_intersection(gmdbuff) %>%
  mutate(COUNTYFP = factor(COUNTYFP))

# # remove Colorado bits and re-factor fips
# gmdCounties <- gmdCounties %>% 
#   filter(abb == 'KS') %>%
#   mutate(fips5 = factor(fips5))

plot(gmdCounties['COUNTYFP'], main = 'Counties In Buffer', lty=0, key.pos=1)
plot(gmd4.ll, col='transparent',add=TRUE, lty=2)
```

## Expand to Get 4 Counties
Actually, belatedly decided to include a little extra area in order to have 4 full counties for validation - so, add the full polygon of Cheyenne County, KS (20023) with the 5 km buffer

```{r addCheyenne}
# extract Cheyenne
chey <- st_as_sf(counties) %>% 
  filter(fips5 == '20023') 

# redo buffered region
gmdAndCountyBuff <- st_as_sf(gmd4.ll) %>% 
  st_union(chey) %>% 
  st_transform(5070) %>%
  st_buffer(5000) %>%
  st_transform(4326)

# get counties for new buffer
gmd.4Counties <- st_as_sf(counties) %>% 
  st_intersection(gmdAndCountyBuff) %>%
  mutate(COUNTYFP = factor(COUNTYFP))

# plot 
plot(gmd.4Counties['COUNTYFP'], main = 'Counties In Buffer', lty=0, key.pos=1)
plot(gmd4.ll, col='transparent',add=TRUE, lty=2)
plot(lema.ll, col= 'transparent',add=TRUE)
```

## Export Boundary
I'm going to make this my boundary to get unique climate-soil-crop-management combos. 

```{r exportBounds}
# get rid of data frame schlep
gmdCountiesOut <- gmd.4Counties[,c('masterID','fips5','NAME','abb','geometry')]

# and make a dissolved border too
boundary <- st_union(gmdCountiesOut)
boundary2 <- st_sf(masterID = 'gmd4_plus', geom = boundary)

# # export
# st_write(gmdCountiesOut, paste0(gisDir,'/GMD4_plus_byCounty.shp'))
# st_write(boundary2, paste0(gisDir,'/GMD4_plus.shp'))
```

### Calculate region areas

```{r areas}
# what's that area?
boundary2 %>% st_transform(5070) %>% st_area() 
# what's that area?
st_as_sf(gmd4.ll)%>% st_transform(5070) %>% st_area() 
#what's that area?
st_as_sf(lema.ll)%>% st_transform(5070) %>% st_area() 
```

