---
title: "NASS County Yields"
author: "Jill Deines"
date: "November 15, 2017"
output: 
  html_document:
    toc: yes
---

Goal: Retrieve crop yields by county from the NASS Quickstats API. 

Functions to package up:

* get yields from NASS - incorporated with the management function, or a separate one (many small, one large?)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path='figure/02.30_salus_nassYields/',
                      cache = TRUE, cache.path='cache/02.30_salus_nassYields/')
```

**R Packages Needed**

```{r packages, warning=FALSE, message=FALSE}
# for API interaction
library(httr) 
library(jsonlite)

# to pull county names
library(sf)

library(tidyverse)
```

**Directories**

```{r dirVars}
# place to store output
yieldDir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/LEMAs/data/salus/nass_yield'

# studay area boundary with counties
boundFile <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/LEMAs/data/GIS/boundaries/GMD4_plus_byCounty.shp'
```


# Functions: NASS Quick Stats API 

* API website: http://quickstats.nass.usda.gov/api

## Function: Get Yield by Commodity, County, Year
part of the set of functions i've made for NASS that really should become 1 or 2.

Current hard coded: 

* sector: 'CROPS'
* group: 'FIELD CROPS' 
* Domain: 'TOTAL'
* Geographic Level: 'STATE'

Function parameters:

* apiKey: Obtained from http://quickstats.nass.usda.gov/api
* program: 'SURVEY' (other option: 'CENSUS')
* Commodity: 'CORN' ('WHEAT', 'SORGHUM','SOYBEANS')
* Category: 'YIELD' 
* Data Item/ short desc: ' - PROGRESS, MEASURED IN PCT PLANTED' ...etc, pasted to crop type but not exactly
* State: 'KS' - 2 letter code 
* county: county 3-digit ANSI code
* Year: '2006' = start  year; retrieves data for all subsequent years

Requires: `httr` and `jsonlite`

```{r fun_getYield}
# # function testing variables
# apiKey <- '28EAA9E6-8060-34A4-981A-B2ED4692228A'
# program <- 'SURVEY'
# commodity <- 'CORN'
# shortDesc <- 'CORN, GRAIN, IRRIGATED - YIELD, MEASURED IN BU / ACRE'
# state <- 'KS'
# county <- 179
# year <- 2006

getCountyYieldNASS <- function(apikey, program, commodity, shortDesc, state, 
                               county, year){
  # build URL query
  baseurl <- 'http://quickstats.nass.usda.gov/api/api_GET'
  format = 'JSON'
  GETrequest <- paste0(baseurl,
                     '/?key=',apiKey,
                     '&source_desc=', program,
                     '&sector_desc=', 'CROPS',
                     '&group_desc=', 'FIELD CROPS',
                     '&commodity_desc=', commodity,
                     '&statisticcat_desc=', 'YIELD',
                     '&short_desc=', shortDesc,
                     '&domain_desc=', 'TOTAL',
                     '&agg_level_desc=', 'COUNTY',
                     '&state_alpha=', state,
                     '&county_ansi=', county,
                     '&year__GE=', year,
                     '&format=', format)
  
  # if present, replace commas and spaces in url with encodings
  if(grepl(" ", GETrequest))  GETrequest <- gsub(" ", "%20", GETrequest)
  if(grepl(",", GETrequest))  GETrequest <- gsub(",", "%2C", GETrequest)
  
  # retrive data
  req <- GET(GETrequest)
  # check status and throw stop/error: 200 means successful
  warn_for_status(req, paste('download', state,county,shortDesc))
  # extract content
  json <- content(req, as = 'text', encoding = 'UTF-8')
  # convert from JSON and extract df from list object
  outputdf <- fromJSON(json)[[1]]
}
```


# Get Yield Data

## Get Counties of Interest
Pull the county FIPS/state from the project boundary shapefile by county

```{r getCounties}
# load 
boundByCounty <- read_sf(boundFile) %>% as.data.frame()

# get state and 3 digit fips
countiesToGet <- boundByCounty %>%
  mutate(fips3 = substr(fips5, 3,5)) %>%
  select(c(abb, fips3))
```

## Query Parameters
Make a query parameter data frame

### short description types
combinations for the data items based on crop type
```{r makeShorts}
# yield data types short description bases: all, irrigated, rainfed
yieldTypes <- paste0(c('', ', IRRIGATED', ', NON-IRRIGATED'), ' - YIELD, MEASURED IN BU / ACRE')

# crop short desc prefixes
shortPrefixes <- c('CORN, GRAIN', 'SORGHUM, GRAIN', 'SOYBEANS','WHEAT')

# full set of short descriptions needed
shortDescs <- sort(paste0(rep(shortPrefixes,3), yieldTypes))

# df of all combos
cropYieldsToGet <- data.frame(crop = c(rep('CORN',3), rep('SORGHUM',3), 
                                       rep('SOYBEANS',3), rep('WHEAT',3)),
                              short = shortDescs)
```

## Retrieve data
behold the nested for lopp

```{r getYield}
# parameters
apiKey <- '28EAA9E6-8060-34A4-981A-B2ED4692228A'
program <- 'SURVEY'
year <- 2006

# list to store output
yield.list <- vector('list',nrow(countiesToGet)*nrow(cropYieldsToGet))

# get data
x <- 1  # start list indexing at 1
for (i in 1:nrow(countiesToGet)){
  # set variables
  state <-  countiesToGet[i,'abb']
  county <- countiesToGet[i,'fips3']
  
  for(m in 1:nrow(cropYieldsToGet)){
    crop <- as.character(cropYieldsToGet[m,'crop'])
    short <- as.character(cropYieldsToGet[m,'short'])
    
    # retrieve data!
    yield.list[[x]] <- getCountyYieldNASS(apikey, program, crop, 
                               short, state, county, year)
    x <- x + 1  # advance list index
  }      
}
```

A quick check of the warnings indicates that for those counties, there's no data for that crop/short description for the years 2006+

### Format Data
make it nice

```{r formatYield}
# remove bad queries from list
yields.list2 <- yields.list[]

```




