---
title: "Generate Salus Experiment Files"
author: "Jill Deines"
date: "November 23, 2017"
output: 
  html_document:
    toc: yes
---

Goal: Make Experiment Files; here, for LEMA, historical, auto irrigation. 

This makes

* 1 experiment .xdb.xml file for every 1000 experiments (so for LEMA historic, 60 xdbs)
* the HPCC shell script (.sh)
* the HPCC batch script (.bat)

Most functions for this are sourced from the `salustools` package

output salus files successfully run on the HPC on 12/2/17 to 12/3/17

Update 12/4/17: re-did for lema_historic_auto_v02, which uses the updated .cdb.xml to fix sorghum and has a start DOY of 1 in 2005. rainfed corn also planted less densely

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path='figure/02.52_salus_expt_lema/',
                      cache = FALSE, cache.path='cache/02.52_salus_expt_lema/')
```

**R Packages Needed**

```{r packages, warning=FALSE, message=FALSE}
library(tidyverse)
library(stringr)
library(sf)
#library(devtools)
#install_github("jdeines/salustools")
library(salustools) # from Jill's github
library(lubridate)
library(viridis)
```

**Directories**
buried below.


## Extract some management info
Combine planting dates, spacing information, cultivars, etc

From Lydia and Brian:  

**Cultivars:**

* maize: complex model, cultivar MZH, speciesID 'MZ'
* wheat: complex model, cultivar IB1015, 'WH'
  * also a simple winter wheat: 'WW'
* soybeans: simple model (no cultivars) - species ID = 'SB'
* sorghum: simple model (no cultivars) - 'SG'
* grassland - a few grasses or generic pasture. 
  * grass-weeds, speciesID ='GW' - possible complex?
  * generic-pasture: 'PA' - definitely a simple. still needs to be planted and harvested since salus doesn't do perennials. using doy 90 and 325
* fallow: just set the planting date for the next crop and let bare soil run. There does also seem to be a Fallow option (speciesID 'FA'). Not sure the best way to make fallow work. I'm going to plant a really sparse pasture.

**Planting Density**  
corn, soy, and wheat planting density taken from Brian and Lydia's scripts.

from https://www.bookstore.ksre.ksu.edu/pubs/l818.pdf - KS planting guide

**sorghum:** dependent on rainfail   

* density: 35,000/acre = 8.64 per m2 (would be lower for western gmd4)
* 65% field emergence
* row spacing: 10 - 30 inches, varying inches between seeds (20 in - 50.8 cm)
* planting for zone 1: may 15-june 10

** alfalfa **
* spacing in central KS (includes Sheridan): 20 - 30 inches
* 220,000 seeds in a pound
* dryland: 10 pounds per acre - 543 seeds per m2
* irrigated:  10-20 pounds per acre - 15 lbs = 815 seeds
* planted in KS zone 1 between April 25 - JUne 1

**soybean:**

* row width: state data says 19.2 - 23.5 (lets go with 54 cm)
* pdf says ... just going with what Brian has used
* 80% emergence

**corn:** different by irrigation systems, more sparse in western kansas (zone1)  

* northwest dryland: 13,000 - 18,000 plants per acre (3.2-4.44 per m2)
* irrigated: 24,000-34,000 (5.9 - 8.4)

**wheat**: nothing helpful  


* from NASS, KS statewide corn row spacing is about 30 in (76.2 cm)
* from NASS, KS soybeans are 19.2-23.5 inches (lets go with  54 cm)


Currently, emergence = planting, as in other salus examples

```{r getManagement}
startDOY <- 1
startYear <- 2006
# adjust start year for winter wheat planting
wheatStart <- startYear -1

endYear <- 2016
Nyrs <- 12

runTitle <- 'Lema_historic_03'


# load date file
managementDir <- 'S:/Users/deinesji/HPA/Salus/lema/management'
dates <- read.csv(paste0(managementDir,'/NASS_dates_planting_maturity_harvest_KS_2006-2017_4crops_MEDIANs.csv'))

# crop parameter key
cropParamKey <- data.frame(crop = c('CORN','SOYBEANS','WHEAT','SORGHUM','GRASS',
                                    'FALLOW', 'ALFALFA'),
                           CropMod = c('C','S','C','S','S','S','S'),
                           SpeciesID = c('MZ','SB','WH','SG','PA','PA', 'AL'),
                           cultivar = c('MZH','','IB1015','','','',''),
                           Ppop_rain = c('3', '39.54', '494.21', '8.6', '50', '5', '543'),
                           Ppop_irr = c('7.5', '39.54', '494.21', '16', '50', '5','815'),
                           RowSpc = c('88.9', '54', '19.05','50.8', '20', '50', '76'),
                          RowSp_irr = c('75', '54', '19.05','40', '20', '50','52'),
                           tillage = c('R','R','N','N','N','N','N'),
                           # GMD4 is in zone 1, adjust state planting dates?
                           plantDOY_adj = c(5,-5,-10,-5, 0, 0,10),
                           # fertilizer management - right now just always the same
                           numFertEvents = c(2,0,1,1,0,0,1),
                           ANFer1=c(rep(70,7)),
                           ANFer2=c(rep(70,7)),
                           # use median harvest date or auto harvest?
                           harvestFlag = c('M','M','M','M','R','R','M'))



#  to add rows for GRASS and FALLOW
grassFallow <- data.frame(year = rep(startYear:endYear,2),
                          crop = c('GRASS','FALLOW'),
                          plantingDOY_50 = 90,
                          harvestedDOY_50 = 325,
                          plantYear = rep(startYear:endYear,2),
                          harvYear = rep(startYear:endYear,2))

alfalfa <- data.frame(year = startYear:endYear,
                          crop = c('ALFALFA'),
                          plantingDOY_50 = 127,
                          harvestedDOY_50 = 182,
                          plantYear = startYear:endYear,
                          harvYear = startYear:endYear)
toAdd <- rbind(alfalfa, grassFallow)

# join and keep key fields
cropManagement00 <- dates %>%
    mutate(plantYear = year(ymd(plantingDate_50)),
           harvYear = year(ymd(harvestDate_50))) %>%
  select(c('year','crop','plantingDOY_50','harvestedDOY_50','plantYear',
           'harvYear')) 

# add fallow and grass
cropManagement <- rbind(cropManagement00, toAdd)  %>%
  left_join(cropParamKey) %>%
  filter( year <= 2016) %>%
  arrange(year, crop)

#cropManagement %>% group_by(crop) %>% summarize(meanPlant = mean(plantingDOY_50))

cropkey <- data.frame(cropCode = c('001','004','005','024','036','061','176'),
                       crop = c('CORN','SORGHUM','SOYBEANS','WHEAT','ALFALFA',
                                    'FALLOW','GRASS'))
```


## Extract Lema scenarios
Load the unique experiments and split into ~ 71 experiment files of 1000 runs each

Note the ExpID's are linked back to the grid cells by the gridExps object also loaded with uniqueExperiments_gmd4p_Sheridan6.RData

```{r getCombos}
# load 'lemaExps' from 2.21
#comboDir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/LEMAs/data/salus/combinationGrids/spatialFilters'

comboDir <- 'S:/Users/deinesji/HPA/Salus/lema/uniqueExperiments'
load(file = paste0(comboDir,'/uniqueExperiments_gmd4p_Sheridan6_top7.RData'))

# how to break down into xml files? by NLDAS or NLDAS + soil?
lemaExps2 <- lemaExps %>% 
  mutate(nldas = str_sub(ExpCode, start = 10, end = 13),
         state = str_sub(ExpCode, start = 1, end = 3),
         mukey = str_sub(ExpCode, start = 3, end = 9),
         static = str_sub(ExpCode, start = 1, end = 13),
         ExpCode = as.character(ExpCode))

# make a grouping vector
ngroup <- ceiling(nrow(lemaExps)/1000)
groups <- sort(rep(1:ngroup, 1000))[1:nrow(lemaExps2)]

# Assign into groups of 1000, sequentially
lemaExps3 <- lemaExps2 %>% 
  mutate(xdbGroup = groups, xdbCode = paste0('X_',xdbGroup))
```

## Write the experiment files!

```{r writeExpts, cache=FALSE}
# output directory
OutDir <- 'E:/Jill/salusExperiments/3_lema_historic_fixed_v01'
cropfile <- 'cropsn29Dec2016_SGttMatrFixed.cdb.xml'

#OutDir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/LEMAs/data/salus/experiments/LEMA_historical_autoIrrigation_v01/xdbs'

sleepDuration <- 0 # not needed on hydrocloud!

irrFlag <- 'F'
fixedIrr <- 25


# write an experiment file for each xdb group
for(g in 1:ngroup){
  
  expt_set <- lemaExps3 %>% filter(xdbGroup == g)
  
  # translate variables in parameters needed to specify experiment tag
  exp_master <- salustools::makeExperimentTable(runTitle, expt_set$ExpID,
                      expt_set$mukey, expt_set$nldas, expt_set$ExpCode,
                      wheatStart, Nyrs, startDOY, cropfp = cropfile)
  
  
  # Write the file! --------------------------------------------------------------
  
  
  outFile2 <- paste0(OutDir,'/',expt_set$xdbCode[1], '.xdb.xml')
  
  # initialize the file
  write_xdb_topMatter(outFile2)
  #Sys.sleep(sleepDuration)
  
  # for each experiment:
  for(i in 1:nrow(exp_master)){  
    # write each set of experiment parameters
    write_xdb_experiment(outFile2, exp_master$ExpID[i], exp_master$runTitle[i], 
                         exp_master$startYear[i], exp_master$Nyrs[i], exp_master$startDOY[i], 
                         exp_master$nldas[i], exp_master$weatherfp[i], exp_master$soilId[i], 
                         exp_master$soilfp[i], exp_master$cropfp[i])
    #Sys.sleep(sleepDuration)
    
    # parse experiment rotations
    rotations <- salustools::parseRotationStrings(as.character(exp_master$ExpCode[i]),
                    startyear = exp_master$startYear[i]+1, cropkey)
    # combine with management table
    annualSpecs <- rotations %>% left_join(cropManagement,by = c("year", "crop"))
    
    # write rotation and management parameters for each year
    for(m in 1:nrow(annualSpecs)){
      # set irrigation flags
      if (annualSpecs$irrStatus[m] == 1) {
        irrig <-irrFlag
        ppopCol <- 'Ppop_irr'
        rowsCol <- 'RowSp_irr'
      }  
      if (annualSpecs$irrStatus[m] == 0){
        irrig <- 'N'
        ppopCol <- 'Ppop_rain'
        rowsCol <- 'RowSpc'
      } 
      
      if (annualSpecs$numFertEvents[m] > 0) fert <- 'R'
      if (annualSpecs$numFertEvents[m] == 0) fert <- 'N'
      
   
      # rotation parameters
      write_xdb_rotation(outFile2, m, annualSpecs$crop[m], irrig, fert, 
                         annualSpecs$tillage[m], annualSpecs$harvestFlag[m])
      #Sys.sleep(sleepDuration)
      
      # management: planting
      
      ## adjust KS state median for nw region
      adjPlantDOY <- annualSpecs$plantingDOY_50[m] + annualSpecs$plantDOY_adj[m]
      
      write_xdb_mPlanting(outFile2, annualSpecs$CropMod[m], annualSpecs$SpeciesID[m], 
                          annualSpecs$cultivar[m], annualSpecs$plantYear[m], 
                          adjPlantDOY, annualSpecs[m,ppopCol],
                          annualSpecs[m,rowsCol])
      #Sys.sleep(sleepDuration)
      # management: fertilize
      if(fert == 'R'){
        # figure out how many events
        for(p in 1:annualSpecs[m,'numFertEvents']){
          # set first fert date based on planting
          fert1Day <- adjPlantDOY
          # but not for wheat
          if(annualSpecs$crop[m] == 'WHEAT') fert1Day <- 75
          
          if(p==1){
            write_xdb_mFertilize(outFile2, annualSpecs$harvYear[m], fert1Day,
                                 annualSpecs$ANFer1[m])
          }
          if(p==2){
            fert2DOY <- adjPlantDOY + 30
            write_xdb_mFertilize(outFile2, annualSpecs$harvYear[m], fert2DOY,
                                 annualSpecs$ANFer2[m])
          }
        }
      }
      #Sys.sleep(sleepDuration)
      
      # management: tillage
      if(annualSpecs$tillage[m] == 'R'){
        write_xdb_mTillage(outFile2, annualSpecs$harvYear[m], adjPlantDOY)
      }
      #Sys.sleep(sleepDuration)
      
      # management: irrigation
      if(irrig == irrFlag){
        write_xdb_mIrrigate_Auto(outFile2, AIrAm = fixedIrr)
      }
      #Sys.sleep(sleepDuration)
      
      # management: harvest
      if(annualSpecs$harvestFlag[m] == 'M'){
        write_xdb_mHarvest_maturity(outFile2, closeComponent = 'Y')
      }  
      if(annualSpecs$harvestFlag[m] == 'R'){
        write_xdb_mHarvest_reported(outFile2, annualSpecs$harvYear[m], 
                      annualSpecs$harvestedDOY_50[m], closeComponent = 'Y')
      }
      #Sys.sleep(sleepDuration)
    }  
    
    # close rotations and experiment
    write_xdb_bottomMatter(outFile2, closeRotation = 'Y', closeExperiment = 'Y')
    #Sys.sleep(sleepDuration)
  }  
  # close out
  write_xdb_bottomMatter(outFile2, writeVersion = 'Y', closeXDB = 'Y')
  #Sys.sleep(sleepDuration)
}
```


## Make the HPCC scripts

### Shell script
make 1 sh file for each experiment

```{r hpccSh, cache=FALSE}
xdbDir <- 'E:/Jill/salusExperiments/3_lema_historic_fixed_v01'
outDir <- xdbDir

hpcHomeDir <- '/mnt/home/deinesji/salus/3_LEMA_hist_fixed_v01/'
hpcOutDir <- '/mnt/home/deinesji/salus/3_LEMA_hist_fixed_v01/results/'

# be mindful of extensions
sdb <- "KS"
wdbZip <- 'nldas_wdb_xmls_gmd4p_cleaned.tar.gz'
cropfile <- 'cropsn29Dec2016_SGttMatrFixed'

# extract xdb name from xmls
xdb <- sapply(strsplit(list.files(xdbDir, pattern='*.xdb.xml'),'.', fixed=TRUE), head, 1)

DayVars <- 'ExpID,Title,SpeciesID,GWAD,IRRC,DRNC,PREC,LAI'
SeaVars <- 'ExpID,Title,SpeciesID,GWAD'

walltime <- '01:00:00'
memory <- '2000mb'

# loop over xdbs
for (i in 1:length(xdb)){
  salustools::write_HPC_shell(outDir, hpcHomeDir, hpcOutDir, xdb[i], sdb, wdbZip, DayVars,
                 SeaVars, walltime, memory, cdb = cropfile)
}  
```

### HPC bash bat script
sends the sh files to queue

```{r hpccbat, cache=FALSE}
shNames <- paste0(xdb,'.sh')
fileOut <- paste0(outDir, '/LEMA_hist_fixed_v01.bat')

# write out .bat file
write_HPC_bat(shNames, fileOut)
```
