---
title: "Generate Salus Experiment Files"
author: "Jill Deines"
date: "November 23, 2017"
output: 
  html_document:
    toc: yes
---

Goal: Make Experiment Files; here, for LEMA, historical, auto irrigation. 

This makes

* 1 experiment .xdb.xml file for every 1000 experiments (so for LEMA historic, 60 xdbs)
* the HPCC shell script (.sh)
* the HPCC batch script (.bat)

Most functions for this are sourced from the `salustools` package

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path='figure/02.52_salus_expt_lema/',
                      cache = TRUE, cache.path='cache/02.52_salus_expt_lema/')
```

**R Packages Needed**

```{r packages, warning=FALSE, message=FALSE}
library(tidyverse)
library(stringr)
library(sf)
library(salustools) # from Jill's github
library(lubridate)
library(viridis)
```

**Directories**
Not all of them are here. It's a fast messy world.

```{r dirs}

```

## Extract some management info
Combine planting dates, spacing information, cultivars, etc

From Lydia and Brian:  

**Cultivars:**

* maize: complex model, cultivar MZH, speciesID 'MZ'
* wheat: complex model, cultivar IB1015, 'WH'
  * also a simple winter wheat: 'WW'
* soybeans: simple model (no cultivars) - species ID = 'SB'
* sorghum: simple model (no cultivars) - 'SG'
* grassland - a few grasses or generic pasture. 
  * grass-weeds, speciesID ='GW' - possible complex?
  * generic-pasture: 'PA' - definitely a simple. still needs to be planted and harvested since salus doesn't do perennials. using doy 90 and 325
* fallow: just set the planting date for the next crop and let bare soil run. There does also seem to be a Fallow option (speciesID 'FA'). Not sure the best way to make fallow work. I'm going to plant a really sparse pasture.

**Planting Density**  
corn, soy, and wheat planting density taken from Brian and Lydia's scripts.

from https://www.bookstore.ksre.ksu.edu/pubs/l818.pdf - KS planting guide

**sorghum:** dependent on rainfail   

* density: 35,000/acre = 8.64 per m2 (would be lower for western gmd4)
* 65% field emergence
* row spacing: 10 - 30 inches, varying inches between seeds (20 in - 50.8 cm)
* planting for zone 1: may 15-june 10

**soybean:**  

* row width: state data says 19.2 - 23.5 (lets go with 54 cm)
* pdf says ... just going with what Brian has used
* 80% emergence

**corn:** different by irrigation systems, more sparse in western kansas (zone1)  

* northwest dryland: 13,000 - 18,000 plants per acre (3.2-4.44 per m2)
* irrigated: 24,000-34,000 (5.9 - 8.4)

**wheat**: nothing helpful  


* from NASS, KS statewide corn row spacing is about 30 in (76.2 cm)
* from NASS, KS soybeans are 19.2-23.5 inches (lets go with  54 cm)


Currently, emergence = planting, as in other salus examples

```{r getManagement}
startYear <- 2006
endYear <- 2016
Nyrs <- 11
startDOY <- 265
runTitle <- 'Lema_historic_01'
# adjust start year for winter wheat planting
wheatStart <- startYear -1

# load date file
managementDir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/LEMAs/data/salus/management'
dates <- read.csv(paste0(managementDir,'/NASS_dates_planting_maturity_harvest_KS_2006-2017_4crops_MEDIANs.csv'))

# crop parameter key
cropParamKey <- data.frame(crop = c('CORN','SOYBEANS','WHEAT','SORGHUM','GRASS',
                                    'FALLOW'),
                           CropMod = c('C','S','C','S','S','S'),
                           SpeciesID = c('MZ','SB','WH','SG','PA','PA'),
                           cultivar = c('MZH','','IB1015','','',''),
                           Ppop_rain = c('3.2', '39.54', '494.21', '8.6', '50', '5'),
                           Ppop_irr = c('7.5', '39.54', '494.21', '8.6', '50', '5'),
                           RowSpc = c('76', '54', '19.05','50.8', '20', '50'),
                           tillage = c('R','R','N','N','N','N'),
                           # GMD4 is in zone 1, adjust state planting dates?
                           plantDOY_adj = c(5,-5,-10,-5, 0, 0),
                           # fertilizer management - right now just always the same
                           numFertEvents = c(2,0,1,0,0,0),
                           ANFer1=c(rep(70,6)),
                           ANFer2=c(rep(70,6)),
                           # use median harvest date or auto harvest?
                           harvestFlag = c('M','M','M','M','R','R'))

#  to add rows for GRASS and FALLOW
grassFallow <- data.frame(year = rep(startYear:endYear,2),
                          crop = c('GRASS','FALLOW'),
                          plantingDOY_50 = 90,
                          harvestedDOY_50 = 325,
                          plantYear = rep(startYear:endYear,2),
                          harvYear = rep(startYear:endYear,2))

# join and keep key fields
cropManagement00 <- dates %>%
    mutate(plantYear = year(ymd(plantingDate_50)),
           harvYear = year(ymd(harvestDate_50))) %>%
  select(c('year','crop','plantingDOY_50','harvestedDOY_50','plantYear',
           'harvYear')) 

# add fallow and grass
cropManagement <- rbind(cropManagement00, grassFallow) %>%
  left_join(cropParamKey) %>%
  filter( year <= 2016) %>%
  arrange(year, crop)

#cropManagement %>% group_by(crop) %>% summarize(meanPlant = mean(plantingDOY_50))

cropkey <- data.frame(cropCode = c('001','004','005','024','061','176'),
                       crop = c('CORN','SORGHUM','SOYBEANS','WHEAT',
                                    'FALLOW','GRASS'))
```


## Extract Lema scenarios
Load the unique experiments and split into ~ 60 experiment files of 1000 runs each

Note teh ExpID's are linked back to the grid cells by the gridExps object also loaded with uniqueExperiments_gmd4p_Sheridan6.RData

```{r getCombos}
# load 'lemaExps' from 2.21
comboDir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/LEMAs/data/salus/combinationGrids/spatialFilters'
load(file = paste0(comboDir,'/uniqueExperiments_gmd4p_Sheridan6.RData'))

# how to break down into xml files? by NLDAS or NLDAS + soil?
lemaExps2 <- lemaExps %>% 
  mutate(nldas = str_sub(ExpCode, start = 10, end = 13),
         state = str_sub(ExpCode, start = 1, end = 3),
         mukey = str_sub(ExpCode, start = 3, end = 9),
         static = str_sub(ExpCode, start = 1, end = 13),
         ExpCode = as.character(ExpCode))

# make a grouping vector
ngroup <- 60
groups <- sort(rep(1:ngroup, 1000))[1:nrow(lemaExps2)]

# Assign into groups of 1000, sequentially
lemaExps3 <- lemaExps2 %>% 
  mutate(xdbGroup = groups, xdbCode = paste0('X_',xdbGroup))
```

## Write the experiment files!

```{r writeExpts}
# output directory
OutDir <- 'C:/Users/deinesji/Documents/Rprojects/lema_SALUS_experiment_files/lema_historic_v1'

#OutDir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/LEMAs/data/salus/experiments/LEMA_historical_autoIrrigation_v01/xdbs'

# write an experiment file for each xdb group
for(g in 1:ngroup){
  
  expt_set <- lemaExps3 %>% filter(xdbGroup == g)
  
  # translate variables in parameters needed to specify experiment tag
  exp_master <- salustools::makeExperimentTable(runTitle, expt_set$ExpID,
                      expt_set$mukey, expt_set$nldas, expt_set$ExpCode,
                      wheatStart, Nyrs, startDOY)
  
  
  # Write the file! --------------------------------------------------------------
  
  
  outFile2 <- paste0(xbdOutDir,'/',expt_set$xdbCode[1], '.xdb.xml')
  
  # initialize the file
  write_xdb_topMatter(outFile2)
  
  # for each experiment:
  for(i in 1:nrow(exp_master)){  
    # write each set of experiment parameters
    write_xdb_experiment(outFile2, exp_master$ExpID[i], exp_master$runTitle[i], 
                         exp_master$startYear[i], exp_master$Nyrs[i], exp_master$startDOY[i], 
                         exp_master$nldas[i], exp_master$weatherfp[i], exp_master$soilId[i], 
                         exp_master$soilfp[i], exp_master$cropfp[i])
    
    # parse experiment rotations
    rotations <- salustools::parseRotationStrings(as.character(exp_master$ExpCode[i]),
                    startyear = exp_master$startYear[i]+1, cropkey)
    # combine with management table
    annualSpecs <- rotations %>% left_join(cropManagement,by = c("year", "crop"))
    
    # write rotation and management parameters for each year
    for(m in 1:nrow(annualSpecs)){
      # only write a rotation component if it's not fallow
      #if (annualSpecs$crop[m] != 'FALLOW'){
        # set irrigation flags
        if (annualSpecs$irrStatus[m] == 1) {
          irrig <- 'A'
          ppopCol <- 'Ppop_irr'
        }  
        if (annualSpecs$irrStatus[m] == 0){
          irrig <- 'N'
          ppopCol <- 'Ppop_rain'
        } 
        
        if (annualSpecs$numFertEvents[m] > 0) fert <- 'R'
        if (annualSpecs$numFertEvents[m] == 0) fert <- 'N'
        
     
        # rotation parameters
        write_xdb_rotation(outFile2, m, annualSpecs$crop[m], irrig, fert, 
                           annualSpecs$tillage[m], annualSpecs$harvestFlag[m])
        # management: planting
        
        ## adjust KS state median for nw region
        adjPlantDOY <- annualSpecs$plantingDOY_50[m] + annualSpecs$plantDOY_adj[m]
        
        write_xdb_mPlanting(outFile2, annualSpecs$CropMod[m], annualSpecs$SpeciesID[m], 
                            annualSpecs$cultivar[m], annualSpecs$plantYear[m], 
                            adjPlantDOY, annualSpecs[m,ppopCol],
                            annualSpecs[m,'RowSpc'])
        
        # management: fertilize
        if(fert == 'R'){
          # figure out how many events
          for(p in 1:annualSpecs[m,'numFertEvents']){
            # set first fert date based on planting
            fert1Day <- adjPlantDOY
            # but not for wheat
            if(annualSpecs$crop[m] == 'WHEAT') fert1Day <- 75
            
            if(p==1){
              write_xdb_mFertilize(outFile2, annualSpecs$harvYear[m], fert1Day,
                                   annualSpecs$ANFer1[m])
            }
            if(p==2){
              fert2DOY <- adjPlantDOY + 30
              write_xdb_mFertilize(outFile2, annualSpecs$harvYear[m], fert2DOY,
                                   annualSpecs$ANFer2[m])
            }
          }
        }
        
        # management: tillage
        if(annualSpecs$tillage[m] == 'R'){
          write_xdb_mTillage(outFile2, annualSpecs$harvYear[m], adjPlantDOY)
        }
        
        # management: irrigation
        if(irrig == 'A'){
          write_xdb_mIrrigate_Auto(outFile2)
        }
        
        # management: harvest
        if(annualSpecs$harvestFlag[m] == 'M'){
          write_xdb_mHarvest_maturity(outFile2, closeComponent = 'Y')
        }  
        if(annualSpecs$harvestFlag[m] == 'R'){
          write_xdb_mHarvest_reported(outFile2, annualSpecs$harvYear[m], 
                        annualSpecs$harvestedDOY_50[m], closeComponent = 'Y')
        }
     # }  
    }  
    
    # close rotations and experiment
    write_xdb_bottomMatter(outFile2, closeRotation = 'Y', closeExperiment = 'Y')
  }  
  # close out
  write_xdb_bottomMatter(outFile2, writeVersion = 'Y', closeXDB = 'Y')
}
```


## Make the HPCC scripts

### Shell script
only one sh file here, for one experiment

```{r hpccSh}
shDir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/LEMAs/data/salus/experiments/LEMA_historical_autoIrrigation_v01'

hpcHomeDir <- '/mnt/home/deinesji/salus/LEMA_historical_autoIrrigation_v01/'
hpcOutDir <- '/mnt/home/deinesji/salus/LEMA_historical_autoIrrigation_v01/results/'

sdb <- "KS"
wdbZip <- 'nldas_wdb_xmls_gmd4p_cleaned.tar.gz'

xdb <- 'X_1'

DayVars <- 'ExpID,Title,SpeciesID,GWAD,IRRC,DRNC,PREC,LAI'
SeaVars <- 'ExpID,Title,SpeciesID,GWAD,IRRC,DRNC,PREC'

walltime <- '01:00:00'
memory <- '2000mb'

salustools::write_HPC_shell(shDir, hpcHomeDir, hpcOutDir, xdb, sdb, wdbZip, DayVars,
                 SeaVars, walltime, memory)
```

### HPC bash bat script
ends the sh files to queue

```{r hpccbat}
shNames <- paste0(xdb,'.sh')
fileOut <- paste0(shDir, '/LEMA_hist_auto_v01.bat')

# write out .bat file
write_HPC_bat(shNames, fileOut)
```
