---
title: "Generate Salus Experiment Files"
author: "Jill Deines"
date: "November 23, 2017"
output: 
  html_document:
    toc: yes
---

Goal: Make Experiment Files; here, a continuous run of wheat

Functions to package up:

* write experiment file!

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path='figure/02.50_salus_expt/',
                      cache = TRUE, cache.path='cache/02.50_salus_expt/')
```

**R Packages Needed**

```{r packages, warning=FALSE, message=FALSE}
library(tidyverse)
```

**Directories**

```{r dirs}
testDir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/LEMAs/SALUS/testGUIrun/continuousWheat'

```

## Extract some management info

```{r getManagement}
# load date file
managementDir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/LEMAs/data/salus/management'
dates <- read.csv(paste0(managementDir,'/NASS_dates_planting_maturity_harvest_KS_2006-2017_4crops_MEDIANs.csv'))

# extract wheat means
wheatDates <- dates %>% filter(crop == 'WHEAT')
summary(wheatDates$plantingDOY_50)
summary(wheatDates$harvestedDOY_50, na.rm = TRUE)
```


# write Xdb Functions
Pieced together from examining the example .xdb.xmls', Lydia Rill and Brian Baer's python function, Lin's example, and SALUS help files

I think I will break this up into tiny buildable functions:

* write_xdb_topMatter
* write_xdb_experiment
* write_xdb_rotation (with argument for first rotation: starts section)
  * write_xdb_mPlanting - write planting parameters for crop rotation
  * write_xdb_mFertilize -
  * write_xdb_mHarvest
  * etc
* write_xdb_bottomMatter

Functions that could be added?

* write_xdb_measured_data - I don't think I have measured data yet? would be entered below experiment at the same level of Rotation Components

## Function 1: Write XDB Top Matter
This function creates the file and add the 2 lines of top matter - not too much to it.

```{r fun1_write_Top}
# # arguments for development
# outFile <- paste0(testDir,'/wheat_continuous.xdb.xml')

write_xdb_topMatter <- function(outFile){
    # write first rows, creates file (no append = true)
  cat('<?xml version="1.0" encoding="utf-8"?>\n<XDB>\n',file=outFile)
}

# # test
# write_xdb_topMatter(outFile)
```

## Function 2: Write XDB Experiment
Writes the Experiment arguments. This section is closed in the `write_xdb_rotation` function.

```{r fun2_write_Expt}
# arguments for development
outFile <- paste0(testDir,'/wheat_continuous.xdb.xml')
ExpID <- 1
RunTitle <- 'wheatTestContinuous'
startYear <- 2006
endYear <- 2016
startDOY <- 1
WthID <- '1001'
WthFile <- '1001.sdb.xml'
SoilID <- 'KS1017570'
SoilFile <- 'KS.sdb.xml'
CropFile <- 'cropsn29Dec2016.cdb.xml'


write_xdb_experiment <- function(outFile, ExpID, RunTitle, startYear, endYear, startDOY,
                      WthID, WthFile,SoilID, SoilFile, CropFile){
  # calculate subarguments
  nyear <- endYear - startYear + 1
  
  # write Experiment details - appends to previous file
  cat(paste0('  <Experiment ExpID="', ExpID, '" Title="', RunTitle, '" NYrs="', 
             nyear, '" SYear="', startYear, '" SOY="', startDOY,
             # some defaults
             '" ISwWat="Y" ISwNit="Y" ISwPho="N" Frop="1" Soilfp="',
             # soil, weather, crop files
             SoilFile, '" SoilID="', SoilID, '" Weatherfp="', WthFile, 
             '" StationID="', WthID, '" Cropfp="', CropFile, 
             # some defaults
             '" NRrepSq="0" MeEvp="R" MeInf="R" kResOrg="3.914E-5" kSloOrg="0.00013699" >\n'),
      file=outFile, append=TRUE)
}

# test
write_xdb_topMatter(outFile)
write_xdb_experiment(outFile, ExpID, RunTitle, startYear, endYear, startDOY,
                     WthID, WthFile, SoilID, SoilFile, CropFile)
```


## Function 3: Write XDB Rotation_components
Writes the Rotation arguments. If orderNum ==1, also includes the `<Rotation_Components>` start tag for rotation.

As always parameter descriptions at http://salusmodel.glg.msu.edu/salus.ddb.xml

```{r fun3_write_rotation}
# arguments for development
outFile <- paste0(testDir,'/wheat_continuous.xdb.xml')
OrderNum <- 1
Title <- 'Wheat'
irrig <- 'N'  # 'N' - not, 'A' - auto, + more options
fertilize <- 'R' # R = reported date, 'A' - auto, 'N' - not
tillage <- 'N' # N = none, 'A' = automatic,, etc
harvest <- 'R' #R - reported, 'M' = at maturity
PlantDOY <- 276 # median
HarvestDOY <- 178 # median

write_xdb_rotation <- function(outFile, OrderNum, Title, irrig, fertilize, tillage,
                               harvest){
  # open rotation components section if position == 'first'
  if(OrderNum == 1){
    cat('    <Rotation_Components>\n', file = outFile, append=TRUE)
  }
  
  # write Component details - appends to previous file
  cat(paste0('      <Component OrderNum="', OrderNum, '" Title="', Title, 
             '" IPltI="R" IIrrI="', irrig, '" IferI="R" IResI="R" ITilI="', tillage,
             '" IHarI="', harvest, '" IEnvI="N">\n'),
      file=outFile, append=TRUE)
}

# test
write_xdb_topMatter(outFile)
write_xdb_experiment(outFile, ExpID, RunTitle, startYear, endYear, startDOY,
                     WthID, WthFile, SoilID, SoilFile, CropFile)
write_xdb_rotation(outFile, OrderNum, Title, irrig, fertilize, tillage, harvest)
```

### Function Rotation Management: Planting
Make a reported (R, manual) function and an automatic 'A' function?

```{r fun3a_mPlanting}
outFile <- paste0(testDir,'/wheat_continuous.xdb.xml')
cropModel <- 'C' # C = complex, S = simple
species <- 'WH' # 'WH' = wheat
cultivar <- 'P25R40'
year <- 2006
PlantDOY <- 276 # median

write_xdb_mPlanting <- function(outFile, cropMod, plantDOY, ){
  
}


  
  
# test
write_xdb_topMatter(outFile)
write_xdb_experiment(outFile, ExpID, RunTitle, startYear, endYear, startDOY,
                     WthID, WthFile, SoilID, SoilFile, CropFile)
write_xdb_rotation(outFile, OrderNum, Title, irrig, fertilize, tillage, harvest)
```



HarvestDOY <- 178 # median



