---
title: "LEMA Historic Auto Analysis"
author: "Jill Deines"
date: "December 3, 2017"
output: 
  html_document:
    toc: yes
---

Goal: Build SALUS results workflow to extract and aggregate desired SALUS results

Also, examine preliminary results from first LEMA historical run (LEMA_Historic_Auto_v01)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path='figure/03.00_salus_results_lema/',
                      cache = TRUE, cache.path='cache/03.00_salus_results_lema/')
```

**R Packages Needed**

```{r packages, warning=FALSE, message=FALSE}
library(tidyverse)
```

**Directories**

```{r dirs}
# raw salus output directory
scratchDir <- 'F:/Users/deinesji/salusRawResults/lema_historic_auto_v01'

# directory for cleaned/summarized output
outDir <- 'S:/Users/deinesji/HPA/Salus/lema/results/lema_hist_auto_v01'

# run name
runname <- 'lema_historic_auto_v01'

startYear <- 2006 # excluding wheat planting, aka first year of harvest
endYear <- 2016
```

# Clean up output: Exploration
60,000 runs makes about 16 GB of output data. Clean and summarize to reduce final data.

## Seasonal Yields
An easy target: seasonal yields! Extract yield for crops run.

```{r yields, eval=FALSE}
# crop codes desired
crops <- c('MZ', 'WH','SB','SG')

# list seasonal results files
seasonFiles <- list.files(scratchDir, pattern = '*seasonal.csv')

# load first set
sdf <- read_csv(paste0(scratchDir, '/', seasonFiles[1])) %>%
  filter(SpeciesID %in% crops) %>%
  select(-c(RID, RcID))
```

## Inquiry: Is Daily Different?
I'd like to just use the seasonal data for irrigation and drainage, but I have a hunch it just reports these values on the harvest DOY. Confirm by checking daily

```{r dailycomp, eval=FALSE}

dailyFiles <- list.files(scratchDir, pattern = '*daily.csv')

ddf <- read_csv(paste0(scratchDir, '/', dailyFiles[1])) 
 # group_by(ExpID, Year, SpeciesID) %>%
  #summarize()
  
# compare values between seasonal and daily
sdf %>% filter(ExpID == 'L-40' & Year == 2015)
ddf %>% filter(ExpID == 'L-40' & Year == 2015 & DOY %in% c(1,254,365))  
```

So the seasonal yield matches, but the seasonal precip (and I assume drainage?) stops at the DOY of harvest. Precipitation and DRNC also seem to be cumulative across the whole run, not just annually, but GWAD is annual

# Clean Output: Mean it!
From the above test, it seems pointless to use the seasonal file - just summarize and aggregate the daily file to get annual values for all variables.

```{r cleanOUtput, eval=FALSE}
# get a list of all filenames in the SALUS run (contained in a folder)
dailyFiles <- list.files(scratchDir, pattern = '*daily.csv')

# load daily files, get annual non-cumulative values, and combine into 1 df
annualResults <- do.call(rbind, lapply(dailyFiles, function(x){
  # load file
  ddf <- read_csv(paste0(scratchDir, '/', x)) 
  
  # make an annual crop table - speciesID where GWAD is max
  cropByYear <- ddf %>%
    group_by(ExpID,Year) %>%
    slice(which.max(GWAD)) %>%
    select(c(ExpID,Year, SpeciesID))
  
  # get maximum values per year
  adf <- ddf %>%
   group_by(ExpID, Year) %>%
   summarize(GWAD = max(GWAD, na.rm=TRUE),
             IRRC_cum = max(IRRC, na.rm=TRUE),
             DRNC_cum = max(DRNC, na.rm=TRUE),
             PREC_cum = max(PREC, na.rm = TRUE),
             LAI = max(LAI, na.rm = TRUE)) %>%
    # add species name back in 
    left_join(cropByYear, by = c('ExpID','Year')) %>%
    # de-cumulate irrc, drnc, prec (turn to annual values)
    mutate(IRRC_mm = IRRC_cum - lag(IRRC_cum, default=0),
           DRNC_mm = DRNC_cum - lag(DRNC_cum, default=0),
           PREC_mm = PREC_cum - lag(PREC_cum, default=0)) %>%
    select(c(ExpID, Year, SpeciesID, GWAD, IRRC_mm, DRNC_mm, PREC_mm, LAI)) %>%
    filter(Year >= startYear)
}))

# write out
write_csv(annualResults, path = paste0(outDir, '/', runname, '.csv'))
# save as rdata: use .rds to be able to assign variable name when reloadig with readRDS
save(annualResults, file = paste0(outDir, '/', runname, '.Rds'))
```

# Check out preliminary results for v01!
I already know I should re-run with my startDOY at 1. Take a quick look at the first set of results for any other glaring omissios to fix.

## Yields

```{r checkResults_yield, fig.width = 5, fig.height = 4.25}
# load formatted results output
aresults <- read_csv(paste0(outDir, '/', runname, '.csv'))

# add an irrigation variable
aresults$irrigation <- 'N'
aresults[aresults$IRRC_mm > 0 ,'irrigation'] <- 'Y'


# get yield statistics by year/crop/irrigation status (not spatially weighted)
# species names
cropKey <- data.frame(SpeciesID = c('MZ', 'WH','SB','SG'),
                      crop = c('CORN','WHEAT','SOYBEANS','SORGHUM'))

yieldStats <- aresults %>%
  filter(SpeciesID %in% cropKey$SpeciesID) %>%
  group_by(Year, SpeciesID, irrigation) %>%
  do(data.frame(t(quantile(.$GWAD, probs = c(.1, .25, .5, .75, .9),na.rm=TRUE)))) %>%
  left_join(cropKey)

# plot
ggplot(yieldStats, aes(x=Year, y = X50., color = irrigation)) +
  geom_ribbon(aes(x = Year, ymin = X25., ymax = X75., fill=irrigation), alpha = .4) +
  geom_line() +
  facet_wrap(~crop) +
  ylab('Yield (kg/ha)') + xlab('') + 
  ggtitle('Median and IQR Yield, Lema Unweighted Experiments') +
  theme_bw() + theme(legend.position = 'bottom')#, legend.title = element_blank())
```

Not bad, but why do I have no sorghum?

### Sorghum Check
I wonder if my choice of model start date is bad for sorghum somehow, since it's harvest date is the latest? examine the daily sequence

```{r sorghumCheck}
# do any seasonal runs have yield (aka daily processing error)
seasonFiles <- list.files(scratchDir, pattern = '*seasonal.csv')
sg_seas <- read_csv(paste0(scratchDir, '/', seasonFiles[1])) %>%
  filter(SpeciesID == 'SG') %>%
  select(-c(RID, RcID))
unique(sg_seas$GWAD) # nope, all 0

# find a daily sorghum run
sorghum_annual <- aresults %>% filter(SpeciesID == 'SG')
# 265 has 3 rotations

#seasonal results for l-265
sg_seas %>% filter(ExpID == 'L-265') # why is the DOY 152?

```

Note the early harvest DOY!

It's auto-harvesting 2 days after planting! Example, 2014 in L-09 (in the log it's clear)

