---
title: "Process Salus Results"
author: "Jill Deines"
date: "December 4, 2017"
output: 
  html_document:
    toc: yes
---

Goal: Build SALUS results workflow to process LEMA output back into a spatial grid

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path='figure/03.20_5_salus_results_process/',
                      cache = FALSE, cache.path='cache/03.20_4_salus_results_lema/')
```

**R Packages Needed**

```{r packages, warning=FALSE, message=FALSE}
# trying to include package::function but the order of these two packages 
# will matter if I forget
library(raster)  
library(tidyverse)
library(rgdal)
#library(devtools)
#install_github("jdeines/salustools")
library(salustools)
```

**Directories and User Variables**

```{r dirs}
# directory for cleaned/summarized to annual values salus output
resultsDir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/LEMAs/data/salus/experiments/5_LEMA_hist_fixed_v03/results'
# subfolder for maps output
subf <- 'maps'

# run name
runname <- '5_lema_historic_fixed_v03'

# time frame
startYear <- 2006 # excluding wheat planting, aka first year of harvest
endYear <- 2016

# load unique experiments and grid cell to Experiment code key
#comboDir <- 'S:/Users/deinesji/HPA/Salus/lema/uniqueExperiments'
comboDir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/LEMAs/data/salus/combinationGrids/spatialFilters'
load(file = paste0(comboDir,'/uniqueExperiments_gmd4p_Sheridan6_top7.RData'))

# gmd4+ template raster grid (based on CDL clipped to study region boundary)
aeaProj <- "+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=23 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"
template <- raster(nrow = 4359, ncol = 7108, crs = aeaProj,
                    xmn = -522870, xmx = -309630, ymn = 1777860, 
                    ymx = 1908630) 
template[] <- 1:ncell(template)
templateVector <- getValues(template)

# aoi
gisDir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/LEMAs/data/GIS/boundaries'
aoi <- readOGR(gisDir, 'Sheridan6_fromKGS_sd6', verbose = FALSE)
aoi <- spTransform(aoi, CRS(aeaProj))
```

# SALUS Maps and Summaries

## Load Data and get Expt Distributions
Results need to be distributed in space, but first, taken a quick glance at the unweighted results for the experiments run

### YIeld

```{r unweightedYieldResults, fig.width = 5, fig.height = 4.25}
# load formatted results output
expt.annual <- read_csv(paste0(resultsDir, '/', runname, '.csv'))

# get yield statistics by year/crop/irrigation status (not spatially weighted)
# species names
cropKey <- data.frame(SpeciesID = c('MZ', 'WH','SB','SG','AL'),
                      crop = c('CORN','WHEAT','SOYBEANS','SORGHUM','ALFALFA'))

uYieldStats <- expt.annual %>%
  filter(SpeciesID %in% cropKey$SpeciesID) %>%
  group_by(Year, SpeciesID, irrigation) %>%
  do(data.frame(t(quantile(.$GWAD, probs = c(.1, .25, .5, .75, .9),na.rm=TRUE)))) %>%
  left_join(cropKey)

# plot
ggplot(uYieldStats, aes(x=Year, y = X50., color = irrigation)) +
  geom_ribbon(aes(x = Year, ymin = X25., ymax = X75., fill=irrigation), alpha = .4) +
  geom_line() +
  facet_wrap(~crop) +
  ylab('Yield (kg/ha)') + xlab('') + 
  ggtitle('Median and IQR Yield, Lema Unweighted Experiments') +
  theme_bw() + theme(legend.position = 'bottom')#, legend.title = element_blank())
```

### Irrigation Water Used

```{r unweightedIrrResults, fig.height = 2.5, fig.width = 4}
# get summary statistics for all irrigated experiments
uIrrStats <- expt.annual %>%
  filter(irrigation == 'Y') %>%
  group_by(Year) %>%
  do(data.frame(t(quantile(.$IRRC_mm, probs = c(.1, .25, .5, .75, .9),na.rm=TRUE))))

# plot
ggplot(uIrrStats, aes(x=Year, y = X50.)) +
  geom_ribbon(aes(x = Year, ymin = X25., ymax = X75.), alpha = .4) +
  geom_line() +
  ylab('Irrigation (mm)') + xlab('') + 
  ggtitle('Median and IQR Irrigation Applied, Lema Unweighted Experiments') +
  theme_bw() 
```

### Precipitation

```{r unweightedPrecip, fig.height = 2.5, fig.width = 4}
uIPrecipStats <- expt.annual %>%
  group_by(Year) %>%
  do(data.frame(t(quantile(.$PREC_mm, probs = c(.1, .25, .5, .75, .9),na.rm=TRUE))))

ggplot(uIPrecipStats, aes(x=Year, y = X50.)) +
  geom_ribbon(aes(x = Year, ymin = X25., ymax = X75.), alpha = .4) +
  geom_line() +
  ylab('Precip (mm)') + xlab('') + 
  ggtitle('Median and IQR Precip, Lema Unweighted Experiments') +
  theme_bw() 
```


## Spatially Distribute
Distribute results based on grid cell ids. Rasters exported to a maps subfolder of the results directory

## function
Function to apply the spatialize function to all years, and export rasters

```{r funs_spatializeAll}
# returns a raster stack for value
spatializeAll <- function(expt.df, variable, yearsWanted, gridkey, template, aoi,
                          writeOut = 'N', outFolder = ''){
                  rasList <- list()
                  for(year in yearsWanted){
                    rasName <- as.character(year)
                    ras <- salustools::spatialize(expt.df, variable, year, 
                                                  gridkey, template,aoi)
                    if(writeOut == 'Y'){
                      fname <- paste0(outFolder, '/',variable, '_',year, '.tif')
                      writeRaster(ras, file = fname)
                    }
                    rasList[[rasName]] <- ras
                  }  
                  rasStack <- stack(rasList, quick = TRUE)
}
```


## Spatial maps

```{r globalVars}
# load formatted results output
expt.annual <- read_csv(paste0(resultsDir, '/', runname, '.csv'))

# create subfolder if needed
outdir <- paste0(resultsDir, '/', subf)
if(!dir.exists(outdir)){
  dir.create(file.path(resultsDir, subf), FALSE)
}

# vector of years wanted
yearsWanted <- startYear:endYear
```

### Irrigation

```{r spIrr, fig.width = 9, fig.height = 5}
irrStack <- spatializeAll(expt.annual, 'IRRC_mm', yearsWanted, gridExps,
                          template, aoi, writeOut = 'N', outFolder = outdir)
spplot(irrStack, main = 'SALUS Irrigation (mm)', maxpixels = 200000)
```

### Precip

```{r spPrecip, fig.width = 9, fig.height = 5}
pcpStack <- spatializeAll(expt.annual, 'PREC_mm', yearsWanted, gridExps,
                          template, aoi, writeOut = 'N', outFolder = outdir)
spplot(pcpStack, main = 'NLDAS Precip (mm)', maxpixels = 200000)
```

### Recharge

```{r spRecharge, fig.width = 9, fig.height = 5}
rchStack <- spatializeAll(expt.annual, 'DRNC_mm', yearsWanted, gridExps,
                          template, aoi, writeOut = 'N', outFolder = outdir)
spplot(rchStack, main = 'SALUS Recharge (mm)', maxpixels = 200000)


#save the stacks
save(irrStack, pcpStack, rchStack, file = paste0(outdir,
                                                 '/All_Irr_Rch_precip_stacks.RData'))
```

### Yield
Break down by crop

```{r spYield, fig.width = 9, fig.height = 5, eval = FALSE}
yield <- 'GWAD'

crops <- c('MZ','WH','SG','SB','AL')

allCrops <- list()
for(m in 1:length(crops)){
  # subset results for crop
  crop <- crops[m]
  cropdf <- expt.annual %>% filter(SpeciesID == crop)
  cropStack <- spatializeAll(cropdf, yield, yearsWanted, gridExps,
                          template, aoi, writeOut = 'N', outFolder = outdir)
  allCrops[[crop]] <- cropStack
}

# save as RData
save(allCrops, file = paste0(outdir, '/AllCrops_YieldMaps.RData'))

# ie corn yield map
cornStack <- allCrops[['MZ']]
spplot(cornStack, main = 'Corn Yield, kg/ha', maxpixels = 200000)
```

```{r yieldMaps, fig.width = 9, fig.height = 5}
load(file = paste0(outdir, '/AllCrops_YieldMaps.RData'))

# ie corn yield map
cornStack <- allCrops[['MZ']]
spplot(cornStack, main = 'Corn Yield, kg/ha', maxpixels = 200000)

# ie wheat yield map
wheatStack <- allCrops[['WH']]
spplot(wheatStack, main = 'wheat Yield, kg/ha', maxpixels = 200000)
```


## some individual year maps

```{r indvlMaps}
spplot(irrStack[['X2006']], main = '2006 salus irrigation')
spplot(irrStack[['X2012']], main = '2012 salus irrigation')


```


## Data Table Synthesis

* precip is averaged over the domain and kept in mm
* irrigation includes total volumetric water, total area, and mean depth
* drainage...
* yields - summed by crop type

```{r toTable, eval=FALSE}

# get total irrigation over time
irrVolStack <- irrStack * 1e-3 * 30 * 30  # volumetric rasters

irrByYear <- data.frame(year = yearsWanted,
                        irr_m3 = cellStats(irrVolStack, stat = 'sum', na.rm=TRUE))
irrByYear$irr_km3 <- irrByYear$irr_m3 * 1e-6

ggplot(irrByYear, aes(x=year, y = irr_km3)) +
  geom_line() + theme_bw() +
  ylab('Total Irrigation, Cubic km') +
  ggtitle('Salus Irrigation over Time')
```



A few individual year maps, and get mean annual precip over time

```{r spPcp2}
spplot(pcpStack[['X2006']], main = '2006 nldas precipitation')
spplot(pcpStack[['X2012']], main = '2012 nldas precipitation')

# get total ppt  over time
pcpByYear <- data.frame(year = yearsWanted,
                        precip_mm = cellStats(pcpStack, stat = 'mean', na.rm=TRUE))

ggplot(pcpByYear, aes(x=year, y = precip_mm)) +
  geom_line() + theme_bw() +
  ylab('Mean Annual Precip (mm)') +
  ggtitle('Precipitation')
```

#### Precip check
that's a bit bizarre, check against direct NLDAS and GRIDMET data

```{r pcheck}
wdir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/LEMAs/data/salus/weather/weatherCheckSheridan'
nldas <- read_csv(paste0(wdir,'/Sheridan5_nldas_meanPrecip_1000.csv')) %>%
  dplyr::select(c(mean, year)) %>%
  mutate(dataset = 'nldas')

gridmet <- read_csv(paste0(wdir,'/Sheridan5_gridmet_meanPrecip_1000.csv')) %>%
  dplyr::select(c(mean, year)) %>%
  mutate(dataset = 'gridmet')

geePrecip <- rbind(nldas, gridmet)

pcpByYear2 <- pcpByYear %>%
  rename(mean = precip_mm) %>%
  mutate(dataset = 'salus')

allpppt <- rbind(geePrecip, pcpByYear2)

ggplot(allpppt, aes(x=year, y = mean, color = dataset)) +
  geom_line() + ylab('mean annual ppt (mm)') +
  ggtitle('NLDAS/SALUS vs GRIDMET') +
  theme_bw()

```



A few individual year maps, and get total recharge  time series. Includes converting mm to m2

```{r spRch2, eval=FALSE}
spplot(rchStack[['X2006']], main = '2006 recharge')
spplot(rchStack[['X2012']], main = '2012 recharge')

# get total rch over time
rchVolStack <- rchStack * 1e-3 * 30 * 30  # volumetric rasters

rchByYear <- data.frame(year = yearsWanted,
                        rch_m3 = cellStats(rchVolStack, stat = 'sum', na.rm=TRUE))
rchByYear$rch_m3 <- rchByYear$rch_m3 * 1e-6

ggplot(rchByYear, aes(x=year, y = rch_m3)) +
  geom_line() + theme_bw() +
  ylab('Total Recharge, Cubic km') +
  ggtitle('Salus Recharge over Time')
```


### Yield

```{r spYield3, fig.width = 9, fig.height = 5, eval=FALSE}
# load formatted results output
expt.annual <- read_csv(paste0(resultsDir, '/', runname, '.csv'))
variable <- 'GWAD'

# vector of years wanted
yearsWanted <- startYear:endYear

# irrigation stack
yieldList <- list()
for(year in yearsWanted){
  rasName <- as.character(year)
  yieldList[[rasName]] <- salustools::spatialize(expt.annual, variable, year, 
                                        gridExps, template,aoi)
}  
yieldStack <- stack(yieldList, quick = TRUE)

spplot(yieldStack, main = 'Yield, kg/ha, run 3', maxpixels = 200000)
```

A few individual year maps, 

```{r spyield2, eval=FALSE}
spplot(yieldStack[['X2006']], main = '2006 yield (k/ha)')
spplot(yieldStack[['X2012']], main = '2012 yield (k/ha)')


```