---
title: "03.50_salus_valid.Rmd"
author: "Jill Deines"
date: "December 7, 2017"
output: 
  html_document:
    toc: yes
---

Goal: Compare SALUS output to real world data

Update 12/9/2017: the "salus optimum run" (50% soil threshold, 25 mm application) is compared to 2 scenarios trying to get cloers to real world irrigation pre-LEMA

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path='figure/03.50_5_salus_valid/',
                      cache = FALSE)
```

**R Packages Needed**

```{r packages, warning=FALSE, message=FALSE}
library(tidyverse)
# library(devtools)
#devtools::install_github("jdeines/salustools")
library(salustools)
```

**Directories and User Variables**

```{r dirs}
# directory for cleaned/summarized to annual values salus output
resultsDir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/LEMAs/data/salus/experiments/5_LEMA_hist_fixed_v03/results'
watervars <- 'WaterVars_meansTotals.csv'
yieldTotals <- 'yields_meansTotals.csv'

# pumping data from Wimas
welldir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/LEMAs/data/wellData/WIMAS_AnthonyCleaned_JillFormatted'
wellsSheridan <- 'WIMAS_Sheridan_1990-2016.csv'

# cdl and irrigated area from gee exports
tabDir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/LEMAs/data/tabular'
cdl_sd6 <- 'CDL_SheridanTrace_2006-2016.csv'
aimName <- 'irrArea_climate_from_AIMRRB.csv'
```


# Combine Water Data
salus water use, wimas pumping, aim area

```{r waterUse}
saluswater <- read_csv(paste0(resultsDir,'/',watervars)) %>%
  mutate(datasource = 'SALUS') 

saluslong <- saluswater %>%
  dplyr::select(c(year, irr_km3, precip_mm, rch_km3, datasource)) %>%
  gather(., key = variable, value = value, irr_km3:rch_km3)

aim <- read_csv(paste0(tabDir,'/', aimName)) %>%
  filter(masterid == 'sheridan6') %>%
  dplyr::select(c(year,irrigated_km2)) %>%
  mutate(datasource = 'AIM') %>%
  gather(., key = variable, value = value, irrigated_km2)

# summary stats wimas
wimas <- read_csv(paste0(welldir, '/', wellsSheridan))
wimasstats <- wimas %>%
  group_by(year) %>%
  summarise(irr_m3 = sum(volume_m3, na.rm=TRUE),
            totalArea_m2 = sum(area_m2, na.rm=TRUE),
            meanDepth_m = mean(depth_m, na.rm=TRUE)) %>%
  mutate(irrigated_km2 = totalArea_m2 * 1e-6,
         irr_km3 = irr_m3 * 1e-6,
         masterid = 'Sheridan6',
         datasource = 'WIMAS') %>%
  dplyr::select(-c(irr_m3, totalArea_m2, masterid)) %>%
  gather(.,key = variable, value = value, meanDepth_m:irr_km3)

# combine
irrigationSummary <- rbind(rbind(saluslong, aim), wimasstats)

# export
write.csv(irrigationSummary, row.names = FALSE,
          file = paste0(resultsDir,'/irrigationSummary_salus_wimas_aim.csv'))
```

## Irrigated Area

```{r irrArea}
irrArea <- irrigationSummary %>% filter(variable == 'irrigated_km2')
ggplot(irrArea, aes(x = year, y = value, color=datasource)) +
  geom_line() + ylim(0, 125) +
  ylab('Irrigation Area (km^2)') + xlab('') +
  theme_bw() + ggtitle('Irrigated Area Comparison')

ggplot(irrArea[irrArea$year > 2005,], aes(x = year, y = value,
                            color=datasource)) +
  geom_line() +# ylim(0, 125) +
  ylab('Irrigation Area (km^2)') + xlab('') +
  #theme_bw() + 
  ggtitle('Irrigated Area Comparison')
```

## Irrigation Water

```{r irrWater}
irrwater <- irrigationSummary %>% filter(variable == 'irr_km3' & year >= 2006)
ggplot(irrwater, aes(x = year, y = value, color=datasource)) +
  geom_line() + ylab('Irrigation Volume (km^3)') + xlab('') +
  theme_bw() + ggtitle('Evidence for excessive water use prior to LEMA')
```

### Compare with real world estimate scenarios

Load scenarios

```{r loadScenarios}
# directory for cleaned/summarized to annual values salus output
resultsDir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/LEMAs/data/salus/experiments/6_LEMA_hist_realEst_v01/results'
watervars_80p <- 'WaterVars_meansTotals.csv'

salus_80p <- read_csv(paste0(resultsDir,'/',watervars_80p)) %>%
  mutate(datasource = 'SALUS_80p') 

salus_80p_long <- salus_80p %>%
  dplyr::select(c(year, irr_km3, precip_mm, rch_km3, datasource)) %>%
  gather(., key = variable, value = value, irr_km3:rch_km3)

# directory for cleaned/summarized to annual values salus output
resultsDir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/LEMAs/data/salus/experiments/7_LEMA_hist_realEst_v02/results'
watervars_50mm <- 'WaterVars_meansTotals.csv'

salus_50mm <- read_csv(paste0(resultsDir,'/',watervars_50mm)) %>%
  mutate(datasource = 'SALUS_50mm') 

salus_50mm_long <- salus_50mm %>%
  dplyr::select(c(year, irr_km3, precip_mm, rch_km3, datasource)) %>%
  gather(., key = variable, value = value, irr_km3:rch_km3)
```

Combine and plot

```{r irrigationScenarios}
irrigationSummary2 <- do.call('rbind',list(irrigationSummary, salus_50mm_long,
                                           salus_80p_long))

irrwater2 <- irrigationSummary2 %>% filter(variable == 'irr_km3' & year >= 2006)
ggplot(irrwater2, aes(x = year, y = value, color=datasource)) +
  geom_line() + ylab('Irrigation Volume (km^3)') + xlab('') +
  theme_bw() + ggtitle('First attempts at Real Estimate')
```

Increasing the application amount by 100% from 25 mm to 50 mm was waaaayyy too much. Using an 80% threshold wasn't bad.

Focus on the 80% soil water threshold

```{r irrScenarios2}
irrwater3 <- irrwater2 %>% filter(datasource != 'SALUS_50mm')
ggplot(irrwater3, aes(x = year, y = value, color=datasource)) +
  geom_line() + ylab('Irrigation Volume (km^3)') + xlab('') +
  #theme_nqblack() + 
  ggtitle('80% SWC threshold not too bad...')


```








cdl <- read_csv(paste0(tabDir, '/', cdl_sd6))
